The promise of community-driven preprints in ecology and evolution

### 2024-05-03 ms/ms.qmd
Fixing grammar in gender section. Some comments to dela with still

Research papers can take long periods of time to become published (see below). However, Early and Mid-Career Researchers (EMCRs) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to accelerate their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, we expect EMCRs to accept and adopt preprints more than colleagues at later career stages. At the same time, gender differences in preprint use and publication outcomes  have been observed in a number of other research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or women). To obtain an idea of an author's gender, we used the R package *gender* (vers. `r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We only used algorithm-assigned names where the gender was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites (we acknowledge that our approach does not capture self-assigned and non-binary genders). Our data on gender has only two missing values – one where the first name of the submitting author was missing and the other one for a collective submission.

Overall, females were less likely to publish preprints on *EcoEvoRxiv* compared to males (females: `r (table(data4$gender)[1] / sum(table(data4$gender))*100)`% compared to males: `r (table(data4$gender)[2] / sum(table(data4$gender))*100)`%), which is likely a reflection of the broader publishing disparities between male and female scientists [@fox2019gender]. There was also strong evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r res[2, "Std. Error"]`, *p* < 0.001, N = `r nrow(count_dat[complete.cases(count_dat),])`). The majority of preprints were submitted by authors who published their first paper in the last 10 years ([@fig-summary]E), with the median year of first publication of submitting authors being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r round(mean(data2$submitting_author_first_publication_year, na.rm = TRUE),0)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`). These patterns support the expectation that ECMRs may value preprints as a way to make their work more visible and disseminate their finding more quickly.

>>>>

Research papers can take long periods of time to become published (see below). However, Early and Mid-Career Researchers (EMCRs) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to accelerate their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, we expect EMCRs to accept and adopt preprints more than colleagues at later career stages. At the same time, gender differences in preprint use and publication outcomes  have been observed in a number of other research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or women). To obtain an idea of an author's gender, we used the R package *gender* (vers. `r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We acknowledge that our approach does not capture self-assigned and non-binary genders. We only used algorithm-assigned names where the gender was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. Our data on gender has only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission.

Overall, women were less likely to publish preprints on *EcoEvoRxiv* compared to men (women: `r (table(data4$gender)[1] / sum(table(data4$gender))*100)`% compared to men: `r (table(data4$gender)[2] / sum(table(data4$gender))*100)`%), which is likely a reflection of the broader publishing disparities between male and female scientists [@fox2019gender]. There was also strong evidence that the number of preprints posted decreases with later career stages (negative binomial glm: year slope = `r res[2, "Estimate"]`, SE: `r res[2, "Std. Error"]`, *p* < 0.001, N = `r nrow(count_dat[complete.cases(count_dat),])`). The majority of preprints were submitted by authors who published their first paper in the last 10 years ([@fig-summary]E), with the median year of first publication of submitting authors being `r median(data2$submitting_author_first_publication_year, na.rm = TRUE)` (mean = `r round(mean(data2$submitting_author_first_publication_year, na.rm = TRUE),0)`; SD = `r sd(data2$submitting_author_first_publication_year, na.rm = TRUE)`). These patterns support the expectation that EMCRs may value preprints as a way to make their work more visible and disseminate their findings more quickly.
